---
title: Homework (3) Day 5 Plots
author: ''
date: '2021-05-20'
slug: []
categories: ['Lecture Notes']
tags: ['Plots','Data Visualization']
---

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```

### Annotations

**Slide 3**
```{r}
presidential <- subset(presidential, start > economics$date[1])

plot6 <- economics %>% ggplot() + 
  geom_line(aes(date, unemploy))

plot(plot6)

```

**Slide 4**
```{r}
plot7 <- plot6 + 
  geom_rect(aes(xmin = start, xmax = end, fill = party),
            ymin = -Inf, ymax = Inf, alpha = .2,
            data = presidential)

plot(plot7)
```

**Slide 5**
```{r}
plot8 <- plot7 + geom_vline( aes(xintercept = as.numeric(start)),
                       data = presidential, 
                       color = 'grey50', alpha = .5)

plot(plot8)
```

**Slide 6**
```{r}
plot9 <- plot8 + geom_text(aes(x = start, y = 2500, label = name),
                           data = presidential, 
                           size = 3, vjust = 0, hjust = 0, nudge_x = 50)

plot(plot9)
```

**Slide 7**
```{r}
plot10 <- plot9 + scale_fill_manual(values = c('blue','red')) +
  xlab('date') + ylab('unemployment')

plot(plot10)
```

**Slide 8**
```{r}
p <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(data = filter(mpg, manufacturer == 'subaru'),
             color = 'orange',
             size = 3) +
  geom_point()

print(p)
```

**Slide 9**
```{r}
p + annotate(geom = 'point', x = 5.5, y = 40, color = 'orange', size = 3) +
    annotate(geom = 'point', x = 5.5, y = 40) +
    annotate(geom = 'text', x = 5.6, y = 40, label = 'subaru', hjust = 'left')
```  

**Slide 10**
```{r}
p + annotate(geom = 'curve', x = 4, y = 35, xend = 2.65, yend = 27,
             curvature = .3, arrow = arrow(length = unit(2, 'mm'))) +
  annotate(geom = 'text', x = 4.1, y = 35, label = 'subaru', hjust = 'left')
```

### Direct Labeling

**Slide 11**
```{r}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point()
```

**Slide 12**
```{r, message = FALSE, warning = FALSE}
library(directlabels)

ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point(show.legend = FALSE) +
  geom_dl(aes(label = class), method = 'smart.grid')
```

**Slide 13**
```{r, message = FALSE}
library(ggforce)

mpg %>% ggplot(aes(displ, hwy)) +
  geom_point() +
  geom_mark_ellipse(aes(label = cyl, group = cyl)) +
  theme_minimal()

```

Theme Options:

`theme_gray(): default, theme_bw(), theme_linedraw(), theme_light, theme_dark(), theme_class(), theme_void()`

**Slide 14**
```{r}
mpg %>% ggplot(aes(displ, hwy)) +
  geom_point() +
  geom_mark_ellipse(aes(label = cyl, group = cyl)) +
  xlim(1, 7.2) + ylim(9,46) +
  theme_minimal()
```
Display Cylinder by Color (Charlie):
```{r}
mpg %>% ggplot(aes(displ, hwy)) +
  geom_point(aes(color = factor(cyl))) +
  geom_mark_ellipse(aes(label = cyl, group = cyl)) +
  xlim(1, 7.2) + ylim(9, 46) +
  labs(color = 'cylinder') +
  theme_minimal()
```

**Slide 15**

Display Cylinder by Color (Dr. Sihm):
```{r}
mpg %>% ggplot(aes(displ, hwy)) +
  geom_point(aes(color = factor(cyl)), show.legend = FALSE) +
  geom_mark_ellipse(aes(label = cyl, group = cyl)) +
  xlim(1,7.2) + ylim(9, 46) +
  theme_minimal()
```

### How to Remove Legend

**Slide 16**

- Remove legend title only

`p + theme(legend.title = element_black())`

- Hide legend for specific geometry only

`p + geom_yOuR_cHoIcE(aes(color = vAr_nAmE), show.legend = FALSE)`

- Hide entire legend

`p + theme(legend.position = 'none')`

**Slide 17**
```{r, message = FALSE, warning = FALSE}
library(gghighlight)
data(Oxboys, package = 'nlme')

Oxboys %>% ggplot(aes(age, height, group = Subject)) +
  geom_line() +
  geom_point() +
  gghighlight(Subject %in% 1:3)
```

### Annotations Across Facets

**Slide 18**
```{r}
diamonds %>% ggplot(aes(log10(carat), log10(price))) +
  geom_bin2d() +
  facet_wrap(~cut, nrow = 1)
```

**Slide 19**

- First, have y-int and slope of trend line over entire dataset
```{r}
linear_model <- lm(log10(price) ~ log10(carat), data = diamonds)
coef_values <- coef(linear_model)
```

- Add this line to all plots across facets. Now we have shared ref. line
```{r}
diamonds %>% ggplot(aes(log10(carat), log10(price))) +
  geom_bin2d() +
  geom_abline(intercept = coef_values[1], slope = coef_values[2], 
              color = 'white', size = .7) +
  facet_wrap(~cut, nrow = 1)
```

**Slide 20**
```{r}
mpg %>% ggplot(aes(displ, hwy, color = factor(cyl))) + 
  geom_point() +
  gghighlight() +
  facet_wrap(~cyl, ncol = 1)
```


### Arranging Plots Side by Side

**Slide 21**

- Create the following 4 plots
```{r}
plot_1 <- mpg %>% ggplot() + geom_point(aes(x = displ, y = hwy))

plot(plot_1)
```

**Slide 22**
```{r}
plot_2 <- mpg %>% ggplot() + geom_bar(aes(x = as.character(year),
                                          fill = drv),
                                      position = 'dodge') + xlab('year')

plot(plot_2)
```

**Slide 23**
```{r}
plot_3 <- mpg %>% ggplot() +
  geom_density(aes(x = hwy, fill = drv), color = NA) +
  facet_wrap(~drv, ncol = 1) # Removing black lines

plot(plot_3)
```

**Slide 24**
```{r}
plot_4 <- mpg %>% ggplot() + stat_summary(aes(drv, hwy, fill = drv),
                                          geom = 'col', fun.data = mean_se) +
  stat_summary(aes(drv, hwy), geom = 'errorbar', fun.data = mean_se, width=.5)

plot(plot_4)
```

**Slide 25**
```{r}
library(patchwork)

plot_1 + plot_2
```

**Slide 26**
```{r}
plot(plot_1 + plot_2)
```

**Slide 27**

- Adding 3 plots will generated 1x3 grid (same default rules as `facet_wrap()`)
```{r}
plot_1 + plot_2 + plot_3
```

**Slide 28**

- Adding 4 plots will generate 2x2
- Notice patchwork handles alignments automatically

```{r}
plot_1 + plot_2 + plot_3 + plot_4
```

**Slide 29**
```{r}
plot_1 + plot_2 + plot_3 + plot_layout(ncol = 2)
```

**Slide 30**
```{r}
plot_1/plot_2
```

**Slide 31**
```{r}
plot_1 | plot_2
```

**Slide 32**
```{r}
plot_3 | (plot_2/ (plot_1 | plot_4))
```

**Slide 33**
```{r}
layout_1 <- '
AAB
C#B
CDD
'

plot_1 + plot_2 + plot_3 + plot_4 + plot_layout(design = layout_1)
```

**Slide 34**

- We can put all legends into one place. Redudant ones will be dropped.
```{r}
plot_1 + plot_2 + plot_3 + plot_4 + plot_layout(design = layout_1,
                                                guides = 'collect')
```

**Slide 35**

- Using `guide_area()`, we can put guides in an empty area
```{r}
plot_1 + plot_2 + plot_3 + guide_area() + plot_layout(ncol = 2, guides = 'collect')
```

**Slide 36**

- Modifying subplots
```{r}
plot_1_2 <- plot_1 + plot_2
plot_1_2[[1]] <- plot_1_2[[1]] + theme_dark()
plot_1_2[[2]] <- plot_1_2[[2]] + theme_classic()
plot_1_2
```

**Slide 37**

- Assigning common theme in patchwork by &
```{r}
plot_1 + plot_4 & theme_minimal() # apply GGplot layers with & sign
```

**Slide 38**

- Assign common y-axis in patchwork by &
```{r}
plot_1 + plot_4 & scale_y_continuous(limits = c(0,45))
```

**Slide 39**

- Add annotation using `plot_annotation()`
```{r}
plot_3_4 <- plot_3 + plot_4 + plot_annotation(
  title = 'A closer look at effect of drive train in cars',
  caption = 'Source: mpg dataset in ggplot2'
)

plot_3_4
```

**Slide 40**
```{r}
plot_3_4 + plot_annotation(theme = theme_gray(base_family = 'mono'))
```

**Slide 41**
```{r}
plot_3_4 & theme_gray(base_family = 'mono')
```

**Slide 42**

- Tagging each plot (common in sci. lit.)

```{r}
plot_1_2_3 <- plot_1 | (plot_2 / plot_3)
plot_1_2_3 + plot_annotation(tag_levels = 'I') # uppercase Roman Numerals
```

**Slide 43**

- Use nesting to define new tagging levels.
```{r}
plot_1_2_3[[2]] <- plot_1_2_3[[2]]+ plot_layout(tag_level = 'new')
plot_1_2_3 + plot_annotation(tag_levels = c('I', 'a')) 
# uppercase rom numeral, then lowercase alphabet
```

**Slide 44**

- Use `inset_element()`
```{r}
plot_1 + inset_element(plot_2, left = .5, bottom = .4,
                       right = .9, top = .95)

# `npc`: Normalized Parent Coordinate
# relative unit 0 to 1
```

**Slide 45**

- Use different unit with `inset_element()`
```{r}
plot_1_2_a <- plot_1 + inset_element(plot_2, left = .5, bottom = .4,
                                     right = unit(1, 'npc') - unit(5, 'mm'),
                                     top = unit(1, 'npc') - unit(3, 'mm'),
                                     align_to = 'full')
plot(plot_1_2_a)
```

**Slide 46**

- Change theme using & likewise
```{r}
plot_1_2_a & theme_minimal()
```

**Slide 47**

- Tagging works as expected
```{r}
plot_1_2_a + plot_annotation(tag_levels = 'A') & theme_dark()
```
