---
title: Dataset for Final Project
author: 'Charlie H.'
date: '2021-06-16'
slug: []
categories: ['final project', 'dcs 550']
tags: []
---

```{r include = FALSE}
library(tidyverse)
library(skimr)
library(knitr)
```

### Loading and Tidying the Dataset

```{r message = FALSE, warning = FALSE}
bicycle <- read_delim('nc_bicycle_crash.csv', delim = ';',
           na = c('#NULL!', ''))
```

#### Selecting and Renaming Columns for Clarity

```{r message = FALSE, warning = FALSE}
bicycle <- bicycle %>%
  transmute(
    
    City = city,
    CrashDate = as.Date(crashdate, '%d%b%Y'),
    
    BikerAgeGroup = as.factor(bikeagegrp),
    DriverAgeGroup = as.factor(drvragegrp),

    BikerAlcohol = as.factor(bikealcoho),
    DriverAlcohol = as.factor(drvralcoho),

    BikerInjury = as.factor(bikeinjury),
    DriverInjury = as.factor(drvrinjury),
    
    DriverVehicleType = as.factor(drvrvehtyp),
    DriverEstimatedSpeed = as.factor(drvrestspd),
    SpeedLimit = as.factor(speedlimit),

    RoadCondition = as.factor(rdconditio),
    LightCondition = as.factor(lightcond),
    
    GeoPts = geo_point_2d
    
  )
```

#### Separating Longitude and Latitude for Mapping

```{r message = FALSE, warning = FALSE}
bicycle <- bicycle %>% 
  separate(GeoPts, sep = ';',
           into = c('Lat', 'Lon'), remove = TRUE)

bicycle$Lon <- as.numeric(bicycle$Lon)
bicycle$Lat <- as.numeric(bicycle$Lat)
```

-----

### Exploring the Dataset

#### With `head()` Function

``` {r message = FALSE, warning = FALSE}
head(bicycle)
```

#### With `summary()` Function
```{r message = FALSE, warning = FALSE}
summary(bicycle)
```

#### With `skim()` Function
```{r message = FALSE, warning = FALSE}
skim(bicycle)
```

#### With `glimpse()` Function
```{r message = FALSE, warning = FALSE}
glimpse(bicycle)
```

-----

### Variables

There were originally 66 columns in the dataset, so I selected those most relevant to my study. I am interested primarily in surveying the effect of the following variables on both driver and bicyclist injury:

1. Age
2. Alcohol
3. Vehicle Type
4. Speed vs. Speed Limit
5. Road Conditions
6. Lighting Conditions

The additional variables of City and Date are included for organizational purposes. Each variable is defined as follows:

#### Age Group

```{r echo = FALSE}
bicycle %>% select(BikerAgeGroup, DriverAgeGroup) %>%
  head() %>%  kable()
```

These two variables indicate the age range of the bicyclist and driver.

#### Alcohol Use

```{r echo = FALSE}
bicycle %>% select(BikerAlcohol, DriverAlcohol) %>%
  head() %>% kable()
```

These two variables indicate whether the bicyclist or driver were under the influence of alcohol at the time of the crash.

#### Vehicle Type

```{r echo = FALSE}
bicycle %>% select(DriverVehicleType) %>% head() %>% kable()
```

This variable indicates the vehicle type of the driver.

#### Speed and Speed Limit

```{r echo = FALSE}
bicycle %>% select(DriverEstimatedSpeed, SpeedLimit) %>%
  head() %>% kable()
```

These two variables are included for comparative analysis. I would like to examine whether breaking the speed limit has any noticeable correlation with crashes.

#### Road Conditions

```{r echo = FALSE}
bicycle %>% select(RoadCondition) %>% head() %>%
  kable()
```

This variable will provide insight into the possible contributing factor of road conditions at the time of the crash.

#### Lighting Conditions

```{r echo = FALSE}
bicycle %>% select(LightCondition) %>% head() %>% kable()
```

This variable indicates the lighting condition at the time of the crash.

#### City, Crash Date, Geographic Information

```{r echo = FALSE}
bicycle %>% select(City, CrashDate, Lat, Lon) %>% head() %>%
  kable()
```

The City and Crash Date variables are for organization. I have included the geographic data for plotting purposes. 